Undirected Graph Exploration with $\Theta(\log\log n)$ Pebbles

### 2016-05-10 presentation/main.tex
slides alpha

\usepackage[utf8]{inputenc}
\usepackage{epigraph}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{xcolor}
\usepackage{amssymb}

@@

\begin{document}
% \tikzexternalize
\maketitle
\begin{frame}
  \frametitle{Motivation}
  \begin{itemize}
    \item Exploration of huge finite graphs with bounded degree by agents
    \item Agents are entities which are located on a vertex and can move along
      an edge
    \item Agents can drop markers on vertices
    \item Agent \emph{explores} a graph by deterministically visiting all of
      its vertices
  \end{itemize}
\end{frame}

\section{Preliminaries}
\begin{frame}
  \frametitle{Huge graphs}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Finite, but huge
      \item Undirected
      \item Indistinguishable vertices
      \item Bounded degree ($\Delta$)
      \item Edges can locally be enumerated as $0,\dots,\Delta-1$
        (Port)
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/portex}}
    \end{center}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Agents}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Agent carries set of distinguishable pebbles
      \item Formalisation as pebble-machines
        \begin{itemize}
          \item $(Q,F,P,m,\delta,\din,\dout,q_{0})$
          \item $P$: set of pebbles
          \item $m$: length of the used memory tape
          \item $\din$: formalisation of agent entering
          \item $\dout$ formalisation of agent leaving
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \begin{itemize}
      \item As an agent moves on a vertex it observes:
        \begin{itemize}
          \item Pebbles it carries
          \item Pebbles lying on the vertex
          \item Degree of vertex
        \end{itemize}
      \item Agent executes a computation on a vertex to determine:
        \begin{itemize}
          \item Port to leave
          \item Pebbles to leave behind
          \item Pebbles to carry with it
        \end{itemize}
    \end{itemize}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Exploration sequence}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Describe movement of agent by relative movements:
        \begin{itemize}
          \item An exploration sequence has the form $e_{1},\dots, e_{n}$
          \item For a vertex $v$ with degree $d_{v}$ which is entered on port
            $p$ is left through port $p + e_{i}\mod d_{v}$
          \item An exploration sequence is called \emph{universial} for a class
            of graphs if following it explores any graph of that class
          \item<2->[$\rightarrow$] Example:
            $\alt<3-3>{\colorbox{green}{1}}{\colorbox{white!0}{1}},
            \alt<4-4>{\colorbox{green}{3}}{\colorbox{white!0}{3}},
            \alt<5-5>{\colorbox{green}{2}}{\colorbox{white!0}{2}},
            \alt<6-6>{\colorbox{green}{2}}{\colorbox{white!0}{2}}$
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/moveex}}
    \end{center}
  \end{columns}
\end{frame}

\section{Graph traversal}
\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{theorem}[Reingold]
    There is an $\mathcal{O}(n)$-space algorithm producing a universal
    exploration sequence for any regular graph on $n$ vertices.
  \end{theorem}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[decorate,decoration={snake,amplitude=0.3cm, segment length=1cm},->]
        (d1) to (d2);
    \end{tikzpicture}
  \end{center}
  \begin{theorem}
    There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
    that moves along a closed walk and either explores the graph or visits at
    least $n$ distinct vertices, for any graph with bounded degree.
  \end{theorem}
\end{frame}
\begin{frame}
  \frametitle{Closed walk}
  \begin{lemma}[Closed walk]
    An agent following an exploration sequence of the form
    $(e_{1},\dots,e_{k})^{\ast}$ moves along a closed walk.
  \end{lemma}
  \begin{itemize}
    \item Examine triples $(v,\ell,i)$:
      \begin{itemize}
        \item $v$: current vertex
        \item $\ell$: port agent entered $v$ through
        \item $i$: next step of the exploration sequence
      \end{itemize}
    \item Deterministically $(v,\ell,i) \leadsto (v',\ell',i')$
    \item If $t_{1} = (v',\ell',i') \leadsto (v,\ell,i)$ and
      $t_{2} = (v'',\ell'',i'') \leadsto (v,\ell,i)$: $t_{1} = t_{2}$
    \item Only finite triples: projection to vertices yields closed walk
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Universal exploration of $3$-regular graphs}
  \begin{lemma}[Periodic exploration sequence (PES-Lemma)]
    There exists an $\mathcal{O}(\log n)$-space algorithm producing a universal
    exploration sequence $(e_{1},\dots,e_{k})^{\ast}$ for any $3$-regular graph
    with at most $n$ vertices.
  \end{lemma}
  \begin{itemize}
    \item Let $\mathcal{M}$ be the Turing machine from Reingold's Theorem
    \item The exploration sequence computed by $\mathcal{M}$ is universial for
      3-regular graphs with at most $n$ vertices
      \alt<1-1>{
        \begin{itemize}
          \item Assume $3$-regular $G_{0}$ with $n_{0} < n$ vertices is not
            explored
          \item Then there is a vertex that is not traversed
          \item Repeatedly apply \parbox{3\baselineskip}{\resizebox{!}
            {3\baselineskip}{\input{tikz/blowup}}}
          \item Until there are $n$ vertices which contradicts Reingold's
            Theorem
        \end{itemize}
      }{
      \item Produce a periodic universal exploration sequence
        $(e_{1},\dots,e_{k})^{\ast}$
        \uncover<2-2>{
          \begin{itemize}
            \item Construct $\mathcal{M}'$ which simulates $\mathcal{M}$ for
              $3n\cdot c$ steps where $c$ is the number of configurations of
              $M$
            \item Output $e = (e_{1},\dots,e_{k})^{\ast}$
            \item Assume $e$ is not universal; then it loops in $V'$ with
              $|V'|<n$
            \item $\mathcal{M}$ has to enter a vertex twice with the same
              configuration
            \item Contradicts universality of the exploration sequence of
              $\mathcal{M}$
          \end{itemize}
        }
      }
  \end{itemize}
  \pause
  \pause
\end{frame}

\begin{frame}
  \frametitle{3-Regularity}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Establish 3-regularity
        \begin{itemize}
          \item Recall: \parbox{3\baselineskip}{\resizebox{!}
            {3\baselineskip}{\input{tikz/blowup}}}
          \item<2-> Transform $v$ to $3\cdot d_{v}$ distinct vertices
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \uncover<3->{\resizebox{\textwidth}{!}{\input{tikz/regsim}}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{itemize}
    \item Graph $G$
    \item Transform to 3-regular $G_{\text{reg}}$
    \item Applying PES-Lemma to $G_{\text{reg}}$ yields:
  \end{itemize}
  \begin{theorem}
    There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
    that moves along a closed walk and either explores the graph or visits at
    least $n$ distinct vertices, for any graph with bounded degree.
  \end{theorem}
\end{frame}

@@

\begin{frame}
  \frametitle{Tape simulation via pebbles}
  \begin{itemize}
    \item Simulate tape of size $m$ by pebbles:
      \begin{itemize}
        \item<3-> Separate into memory blocks of size $m_{1}$
        \item<4-> Represent every block by one of
          $p_{0},\dots,p_{\frac{m}{m_{1}}-1}$ pebbles
      \end{itemize}
  \end{itemize}
  \uncover<2->{\input{tikz/tapesimint}}
\end{frame}

\begin{frame}
  \frametitle{Tape simulation via pebbles}
  \input{tikz/tapesimpebsep}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Every pebbles needs $2^{m_{1}}$ distinct \enquote{states}
        \begin{itemize}
          \item Placement on a walk with at least $2^{m_{1}}$ distinct vertices
            yields \enquote{states}
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/tapesimwalk}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Mechanisms of simulation}
  \begin{itemize}
    \item Simulation of $M_{\text{Walk}}$ which explores at least $2^{m_{1}}$
      distinct vertices
    \item Identify \emph{distinct} vertices of the walk
    \item Read from and write to simulated tape
    \item Preserve tape content through steps of simulated agent
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Simulation of $M_{\text{Walk}}$}
  \begin{itemize}
    \item $M_{\text{Walk}}$ can be implemented as
      $(\mathcal{O}(1), 0, \mathcal{O}(m_{1}))$ pebble machine
    \item Used variables/pebbles:
      \begin{itemize}
        \item $\Twalk$: stores $M_{\text{Walk}}$'s tape
        \item $\Tsteps$: stores number of taken steps
        \item $\Tid$: stores number of visited distinct vertices
        \item $p_{\text{start}}$: marks the beginning
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Finding distinct vertices}
  Problem: Indistinguishable vertices $\leadsto$ next vertex \enquote{new}?
  \begin{columns}
    \column{0.5\textwidth}
    Solution:
    \begin{enumerate}
      \item Execute step
      \item Drop pebble $p_{\text{tmp}}$
      \item Store $\Twalk$ in $T'_{\text{walk}}$
      \item Restart $M_{\text{Walk}}$
      \item Execute steps until $p_{\text{tmp}}$ is observed
      \item If $\Twalk = T'_{\text{walk}}$ the vertex is new
      \item Otherwise start over
    \end{enumerate}
    \column{0.5\textwidth}
    \input{tikz/disvert}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Memory management}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item $\Thead$ stores head position
      \item $p_{\lfloor\frac{\Thead}{m_{1}}\rfloor}$: pebble of memory block
      \item $\mathit{off} = \Thead\mod m_{1}$: offset in memory block
      \item Reading:
        \begin{itemize}
          \item Retrieve encoding pebble
          \item Return bit with offset
        \end{itemize}
      \item Writing:
        \begin{itemize}
          \item Read bit
          \item On bit-flip move encoding pebble $2^{\mathit{off}}$ bits up or
            down
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/tapesimpebsep}}
    \resizebox{\textwidth}{!}{\input{tikz/tapesimwalk}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Memory management: Writing}
  \begin{itemize}
    \item Read bit
    \item On bit-flip: move pebble
  \end{itemize}
  \resizebox{\textwidth}{!}{\input{tikz/bitwrite}}
\end{frame}

\begin{frame}
  \frametitle{Transfer tape-content through movements}
  \begin{itemize}
    \item Simulated machine moves
    \item Preserve tape-content:
      \begin{itemize}
        \item Drop $p_{\text{next}}$ on next vertex
        \item Transfer encoding pebbles onto walk $\omega'$ starting from
          next vertex
        \item Preserve \enquote{state} by placing at same index
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Recall: Mechanisms of simulation}
  \begin{itemize}
    \item Simulation of $M_{\text{Walk}}$ which explores at least $2^{m_{1}}$
      distinct vertices \uncover<2->{\checkmark}
    \item Identify \emph{distinct} vertices of the walk
      \uncover<3->{\checkmark}
    \item Read from and write to simulated tape \uncover<4->{\checkmark}
    \item Preserve tape content through steps of simulated agent
      \uncover<5->{\checkmark}
  \end{itemize}
  \uncover<6->{
    \begin{theorem}
      There is a constant $c\in\mathbb{N}$, such that for any Graph $G$
      with bounded degree and any $(s,p,2m)$ pebble machine $\mathcal{M}$,
      there exists a $(cs,p+c,m)$ pebble machine $\mathcal{M}'$ that simulates
      the walk of $\mathcal{M}$ or explores $G$.
    \end{theorem}
  }
\end{frame}

\begin{frame}
  \frametitle{Exploring graphs with $\log\log n$ pebbles}
  \begin{itemize}
    \item There is a $(c,0,c\cdot\log n)$ pebble machine which explores
      a graph $G$ with $n$ vertices
    \item Applying above simulation $\log \log n$ times yields
      $(c'^{\log\log n}\cdot c, c'\cdot\log\log n, c)$ pebble machine exploring
      $G$
      \begin{theorem}
        Any bounded-degree graph on at most $n$ vertices can be explored using
        $\mathcal{O}(\log\log n)$ pebbles and memory.
      \end{theorem}
  \end{itemize}
\end{frame}

\section{Limitations of pebble machines}
\begin{frame}
  \frametitle{Limitations of pebbles}
  \begin{itemize}
    \item Construct 3-regular graphs which cannot be traversed by an agent
      $\mathcal{A}$ with $p$ pebbles
    \item Such graphs are called $p$-barriers
      \begin{center}
        \resizebox{0.6\textwidth}{!}{\input{tikz/barrierint}}
      \end{center}
    \item $\mathcal{A}$ cannot traverse a barrier from $v_{1}$ or $v_{2}$
      to $v_{3}$ or $v_{4}$ or vice versa using at most $p$ pebbles
    \item Inductive construction of barriers
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Locality gadget}
  \begin{itemize}
    \item Replace edge by a $(p-1)$-barrier
    \item Enforce \enquote{locality} of pebbles
    \item An agent cannot traverse any gadget without \emph{all} his pebbles
    \item All pebbles of the agent are either carried or placed within the
      adjacent gadgets
      \begin{center}
        \uncover<2->{\resizebox{0.8\textwidth}{!}{\input{tikz/locgadget}}}
      \end{center}
    \item<3-> Original vertices are called \emph{macro} vertices
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Barrier construction: Initial thoughts}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Consider a subset of pebbles of cardinality $r$
      \item Use the locality gadget as edges
      \item Since pebbles are local: enumerate configurations in macro vertex
        $C_{1},\dots,C_{\alpha}$
      \item Project $A$ to a agent without pebbles $B$
      \item $B$ has $\alpha$ states and traverses through macro vertices
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/projection}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Barrier construction}
  \begin{itemize}
    \item $B$ has no pebbles and is therefore prone to a $0$-barrier
    \item Apply locality gadget to all edges
    \item This yields a $p$-barrier for the chosen subset of pebbles
    \item Iterate $p$-barriers for all possible subsets
      \begin{center}
        \resizebox{0.7\textwidth}{!}{\input{tikz/iterbarrier}}
      \end{center}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Barrier base case}
  \begin{theorem}
    For any $q$ non cooperative $s$-state agents without pebbles, there
    exists a 3-regular graph $G$ on $\mathcal{O}(qs)$ vertices with the
    following property: There are two edges $\left\{v_{1},v_{2}\right\}$
    and $\left\{v_{3},v_{4}\right\}$ in $G$, the first labeled 0, such
    that any of the $q$ agents starting in $v_{1}$ or $v_{2}$ does not
    traverse the edge $\left\{v_{3},v_{4}\right\}$.
  \end{theorem}
  \begin{itemize}
    \item For agent $\mathcal{A}$ consider
      $S = \left\{\mathcal{A}_{q}\mid q\in Q\right\}$
    \item Since $|S| = |Q| = q$ there is a 0-barrier with $\mathcal{O}(s^{2})$
      vertices
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Barrier size}
  \begin{itemize}
    \item Iterative construction
  \end{itemize}
  \begin{center}
    \resizebox{0.8\textwidth}{!}{\input{tikz/barriersize}}
  \end{center}
  \begin{theorem}
    For $r\leq p$ and $s\geq 2^{p}$, the number of vertices of $r$-barrier
    $B_{r}$ for the $s$-state agent $\mathcal{A}$ with $p$ pebbles is bounded
    by $\mathcal{O}(s^{8^{r+1}})$.
  \end{theorem}
\end{frame}

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage{epigraph}
\usepackage{tikz}
\usepackage{csquotes}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{mdframed}

@@

\begin{document}
% \tikzexternalize
\maketitle
\begin{frame}
  \frametitle{Motivation}
  \begin{itemize}
    \item Exploration of huge finite graphs by agents
    \item Agents are located on a vertex and move along edges
    \item Agents can drop pebbles on vertices
    \item Agent \emph{explores} a graph by deterministically visiting all of
      its vertices
  \end{itemize}
\end{frame}

\section{Preliminaries}
\begin{frame}
  \frametitle{Huge graphs}
  \begin{columns}
    \column{0.55\textwidth}
    \begin{itemize}
      \item Finite, but huge
      \item Undirected
      \item Indistinguishable vertices:
      \parbox{\baselineskip}{
        \resizebox{!}{\baselineskip}{
          \begin{tikzpicture}
            \node[node] {};
          \end{tikzpicture}
        }
      }
      \item Bounded degree ($\Delta$)
      \item Edges can locally be labeled with
        \textcolor{green}{$0,\dots,\Delta-1$} (Port)
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/portex}}
    \end{center}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Agents}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Agent carries set of distinguishable pebbles
      \item Formalisation as pebble-machines
        \begin{itemize}
          \item $(Q,F,P,m,\delta,\din,\dout,q_{0})$
          \item<2-> $\din$: moving onto vertex
          \item<7-> $\dout$: leaving vertex
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \begin{itemize}
      \item<3-> As an agent moves onto a vertex it observes:
        \begin{itemize}
          \item<4-> Pebbles it carries
          \item<5-> Pebbles on the vertex
          \item<6-> Degree of vertex
        \end{itemize}
      \item<8-> Computation on a vertex determines:
        \begin{itemize}
          \item<9-> Port to leave
          \item<10-> Pebbles to drop
          \item<11-> Pebbles to carry along
        \end{itemize}
    \end{itemize}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Exploration sequence}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Exploration sequence $e_{1},\dots, e_{n}$
      \item Relative movements of agent
      \item Leaving port: $\ell_{i} + e_{i}\mod d_{v}$
      \item \emph{Universial} for a class of graphs if it explores any graph of
        that class
      \item<2->[$\rightarrow$] Example:
        $\alt<3-3>{\colorbox{green}{1}}{\colorbox{white!0}{1}},
        \alt<4-4>{\colorbox{green}{3}}{\colorbox{white!0}{3}},
        \alt<5-5>{\colorbox{green}{2}}{\colorbox{white!0}{2}},
        \alt<6-6>{\colorbox{green}{2}}{\colorbox{white!0}{2}}$
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/moveex}}
    \end{center}
  \end{columns}
\end{frame}

\section{Graph traversal}
\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{theorem}[Reingold]
    There is an $\mathcal{O}(n)$-space algorithm producing a universal
    exploration sequence for any regular graph on $n$ vertices.
  \end{theorem}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
      \uncover<2->{
        \node[rectangle, draw=green, below right=0.3 and 1 of d1]
        {3-Regularity};
      }
      \uncover<3->{
        \node[rectangle, draw=blue, below right=1 and 1.2 of d1]
        {Periodic Universal lemma};
      }
      \uncover<4->{
        \node[rectangle, draw=yellow, below right=1.7 and 1 of d1]
        {Closed walk lemma};
      }
    \end{tikzpicture}
  \end{center}
  \begin{theorem}
    There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
    that moves along a closed walk and either explores the graph or visits at
    least $n$ distinct vertices, for any graph with bounded degree.
  \end{theorem}
\end{frame}

\begin{frame}
  \frametitle{Closed walk}
  \begin{mdframed}[linecolor=yellow]
    \begin{lemma}[Closed walk]
      An agent following an exploration sequence of the form
      $(e_{1},\dots,e_{k})^{\ast}$ moves along a closed walk.
    \end{lemma}
  \end{mdframed}
  \uncover<2->{
    \begin{mdframed}[linecolor=blue]
      \begin{lemma}[Periodic exploration sequence (PES-Lemma)]
        There exists an $\mathcal{O}(\log n)$-space algorithm producing a
        universal exploration sequence $(e_{1},\dots,e_{k})^{\ast}$ for any
        $3$-regular graph with at most $n$ vertices.
      \end{lemma}
    \end{mdframed}
  }
\end{frame}

\begin{frame}
  \frametitle{3-Regularity}
  \begin{mdframed}[linecolor=green]
    Establish 3-regularity
  \end{mdframed}
  \begin{center}
    \uncover<2->{\resizebox{0.7\textwidth}{!}{\input{tikz/regsim}}}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{theorem}[Reingold]
    There is an $\mathcal{O}(n)$-space algorithm producing a universal
    exploration sequence for any regular graph on $n$ vertices.
  \end{theorem}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
      \node[rectangle, draw=green, below right=0.3 and 1 of d1]
        {3-Regularity};
      \node[rectangle, draw=blue, below right=1 and 1.2 of d1]
        {Periodic Universal lemma};
      \node[rectangle, draw=yellow, below right=1.7 and 1 of d1]
        {Closed walk lemma};
    \end{tikzpicture}
  \end{center}
  \begin{theorem}
    There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
    that moves along a closed walk and either explores the graph or visits at
    least $n$ distinct vertices, for any graph with bounded degree.
  \end{theorem}
\end{frame}

@@

\begin{frame}
  \frametitle{Tape simulation via pebbles}
  \begin{itemize}
    \item Simulate $m$ bits by pebbles:
      \begin{itemize}
        \item<3-> Separate into $m_{1}$ sized blocks
        \item<4-> Represent blocks by pebbles
          $p_{0},\dots,p_{\frac{m}{m_{1}}-1}$
      \end{itemize}
  \end{itemize}
  \uncover<2->{\input{tikz/tapesimint}}
\end{frame}

\begin{frame}
  \frametitle{Tape simulation via pebbles}
  \input{tikz/tapesimpebsep}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Pebble needs $2^{m_{1}}$ distinct \enquote{states}
      \item State as index of closed walk
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/tapesimwalk}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Mechanisms of simulation}
  \begin{enumerate}
    \item Simulation of $M_{\text{walk}}$ providing walk with $2^{m_{1}}$
      distinct vertices
    \item Identify \emph{distinct} vertices of walk
    \item Read from and write to simulated tape
    \item Preserve tape content through steps of simulated agent
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{1. Simulation of $M_{\text{walk}}$}
  \begin{itemize}
    \item $M_{\text{Walk}}$ implemented as
      $(\mathcal{O}(1), 0, \mathcal{O}(m_{1}))$ pebble machine
    \item Used variables/pebbles:
      \begin{itemize}
        \item $\Twalk$: stores $M_{\text{walk}}$'s tape
        \item $\Tsteps$: stores number of taken steps
        \item $\Tid$: stores number of visited distinct vertices
        \item $p_{\text{start}}$: marks beginning of walk
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{2. Finding distinct vertices}
  Problem: Indistinguishable vertices $\leadsto$ next vertex \enquote{new}?
  \begin{columns}
    \column{0.5\textwidth}
    Solution:
    \begin{enumerate}
      \item \textcolor<3,8>{green}{Execute step}
      \item \textcolor<4,9>{green}{Drop pebble $p_{\text{tmp}}$}
      \item \textcolor<4,9>{green}{Store $\Twalk$ in $T'_{\text{walk}}$}
      \item \textcolor<5,10>{green}{Restart $M_{\text{walk}}$}
      \item \textcolor<6,11-18>{green}{Execute steps until $p_{\text{tmp}}$ is
        observed}
      \item \textcolor<19-20>{green}{If $\Twalk = T'_{\text{walk}}$ the vertex
        is new (increase $\Tid$)}
      \item \textcolor<7>{green}{Otherwise start over}
    \end{enumerate}
    \column{0.5\textwidth}
    \input{tikz/disvert}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{3. Memory management}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item $\Thead$: head position
      \item $p_{\lfloor\frac{\Thead}{m_{1}}\rfloor}$: pebble of memory block
      \item $\mathit{off} = \Thead\mod m_{1}$: offset in memory block
      \item Reading:
        \begin{itemize}
          \item Retrieve encoding pebble
          \item Return bit with offset
        \end{itemize}
      \item Writing:
        \begin{itemize}
          \item Read bit
          \item On bit-flip move encoding pebble $2^{\mathit{off}}$ bits up or
            down
        \end{itemize}
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/tapesimpebsep}}
    \resizebox{\textwidth}{!}{\input{tikz/tapesimwalk}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{3. Memory management: Writing}
  \begin{itemize}
    \item Read bit
    \item On bit-flip: move pebble
  \end{itemize}
  \resizebox{\textwidth}{!}{\input{tikz/bitwrite}}
\end{frame}

\begin{frame}
  \frametitle{4. Transfer tape-content through movements}
  \begin{itemize}
    \item Simulated agent moves
    \item Preserve tape-content:
      \begin{itemize}
        \item Drop $p_{\text{next}}$ on next vertex $v$
        \item $\omega'$ walk of $M_{\text{walk}}$ starting from $v$
        \item Transfer encoding pebbles onto $\omega'$
        \item Placing pebbles at same index
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Recall: Mechanisms of simulation}
  \begin{itemize}
    \item Simulation of $M_{\text{Walk}}$ which explores at least $2^{m_{1}}$
      distinct vertices \uncover<2->{\textcolor{green}{{\checkmark}}}
    \item Identify \emph{distinct} vertices of the walk
      \uncover<3->{\textcolor{green}{{\checkmark}}}
    \item Read from and write to simulated tape \uncover<4->{\textcolor{green}{{\checkmark}}}
    \item Preserve tape content through steps of simulated agent
      \uncover<5->{\textcolor{green}{{\checkmark}}}
  \end{itemize}
  \uncover<6->{
    \begin{theorem}
      There is a constant $c\in\mathbb{N}$, such that for any Graph $G$
      with bounded degree and any $(s,p,2m)$ pebble machine $\mathcal{M}$,
      there exists a $(cs,p+c,m)$ pebble machine $\mathcal{M}'$ that simulates
      the walk of $\mathcal{M}$ or explores $G$.
    \end{theorem}
  }
\end{frame}

\begin{frame}
  \frametitle{Exploring graphs with $\log\log n$ pebbles}
  \begin{theorem}
    There is a constant $c\in\mathbb{N}$, such that for any Graph $G$
    with bounded degree and any $(s,p,2m)$ pebble machine $\mathcal{M}$,
    there exists a $(cs,p+c,m)$ pebble machine $\mathcal{M}'$ that simulates
    the walk of $\mathcal{M}$ or explores $G$.
  \end{theorem}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
      \node[below right=0.8 and 1 of d1, rectangle,draw=black] {Apply simulation $\log\log n$ times};
    \end{tikzpicture}
  \end{center}
  \begin{theorem}
    Any bounded-degree graph on at most $n$ vertices can be explored using
    $\mathcal{O}(\log\log n)$ pebbles and memory.
  \end{theorem}
\end{frame}

\section{Limitations of pebble machines}
\begin{frame}
  \frametitle{Limitations of pebbles}
  \begin{itemize}
    \item Construct 3-regular graphs agent $\mathcal{A}$ with $p$ pebbles
      cannot traverse
    \item Called $p$-barriers
      \begin{center}
        \resizebox{0.6\textwidth}{!}{\input{tikz/barrierint}}
      \end{center}
    \item $\mathcal{A}$ cannot traverse a $p$-barrier from $v_{1}$ or $v_{2}$
      to $v_{3}$ or $v_{4}$ or vice versa using at most $p$ pebbles
    \item Inductive construction of barriers
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Locality gadget}
  \begin{itemize}
    \item Replace edge by $(p-1)$-barrier
    \item Enforce \enquote{locality} of pebbles
      \begin{center}
        \uncover<2->{\resizebox{0.8\textwidth}{!}{\input{tikz/locgadget}}}
      \end{center}
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Barrier construction: Initial thoughts}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Subset of pebbles of cardinality $r$
      \item Locality gadget as edges
      \item Pebbles are local: configurations in macro vertex
        $C_{1},\dots,C_{\alpha}$
      \item Project $A$ to agent without pebbles $B$
      \item $B$ has $\alpha$ states and traverses macro vertices
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/projection}}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Barrier construction}
  \begin{itemize}
    \item $B$ has no pebbles and cannot traverse $0$-barrier
    \item[$\rightarrow$] $p$-barrier for the chosen subset of pebbles:
  \end{itemize}
  \begin{center}
    \resizebox{0.8\textwidth}{!}{\input{tikz/iterbarrier}}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Barrier base case}
  \begin{theorem}[Fraigniaud et al.]
    For any $q$ non cooperative $s$-state agents without pebbles, there
    exists a 3-regular graph $G$ on $\mathcal{O}(qs)$ vertices with the
    following property: There are two edges $\left\{v_{1},v_{2}\right\}$
    and $\left\{v_{3},v_{4}\right\}$ in $G$, the first labeled 0, such
    that any of the $q$ agents starting in $v_{1}$ or $v_{2}$ does not
    traverse the edge $\left\{v_{3},v_{4}\right\}$.
  \end{theorem}
  \begin{itemize}
    \item For agent $\mathcal{A}$ consider
      $S = \left\{\mathcal{A}_{q}\mid q\in Q\right\}$
    \item Since $|S| = |Q| = q$ there is a 0-barrier with $\mathcal{O}(s^{2})$
      vertices
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Barrier size}
  \begin{center}
    \resizebox{\textwidth}{!}{\input{tikz/barriersize}}
  \end{center}
  \begin{theorem}
    For $r\leq p$ and $s\geq 2^{p}$, the number of vertices of $r$-barrier
    $B_{r}$ for the $s$-state agent $\mathcal{A}$ with $p$ pebbles is bounded
    by $\mathcal{O}(s^{8^{r+1}})$.
  \end{theorem}
\end{frame}

@@

\section{Conclusion}
\begin{frame}
  \frametitle{Conclusion}
  \begin{itemize}
    \item Exploring graphs with $\mathcal{O}(\log\log n)$ pebbles and memory
    \item Simulation via pebbles
    \item Limitations of pebbles: Barrier construction
  \end{itemize}
\end{frame}
